White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification

### 2015-09-22 papers/wdmerger_I/paper.tex
Updated scaling results

To examine the parallel performance of \castro, we performed both
strong scaling and weak scaling tests on the Blue Waters machine at
the National Center for Supercomputing Applications. For the weak
scaling test, whose results are shown in \autoref{fig:weak_scaling},
we ran a uniform grid binary white dwarf simulation for resolutions of
$128^3$ zones through $2048^3$ zones. The number of processors was
scaled with the number of zones so that each processor had the same
amount of work; the smallest test used 8 processors and the largest
used 32,768 (note that the number of processor cores on a Blue Waters
node is twice the number of floating point units on that node). The
test was run for 10 timesteps, with each timestep including two
Poisson solves and a hydrodynamics update (though for a uniform grid
calculation we generally do not need to perform any multigrid
iterations for the first Poisson solve in a timestep, since the
density distribution has not changed since the end of the last
timestep). We disabled plotfile and checkpoint writing, as well as
calculation of diagnostic information (the latter can contribute to a
significant fraction of the run time at large processor counts if
computed every timestep). We computed the median wall time required
per time step for each simulation, and then normalized this to the
median time per timestep for the smallest simulation. We find
excellent weak scaling through 4,096 processors. At the largest run,
the simulation time required is more than 1.5 times the amount
required for the smallest simulation.  This is due entirely to the
increased cost of the multigrid Poisson solve in each timestep and
this cannot be mitigated except by improving communication or
computation efficiency in the multigrid solver. We observe that this
weak scaling behavior with Poisson gravity is a significant
improvement over the results presented in the first \castro\ paper.

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{plots/strong_scaling}
  \caption{\castro\ strong scaling test performed on the Blue Waters machine at
    NCSA. The vertical axis measures the median time per timestep, and the 
    horizontal axis measures the number of processors in the simulation. Data 
    points are normalized to the time per timestep for the smallest number 
    of processors. The blue circles show the data for a simulation with two AMR
    levels (one coarse and one fine), while the red circles show the data 
    for a test with three AMR levels (one coarse and two fine). The fine levels 
    increase the resolution only in the regions around the stars. For each case 
    we draw a solid curve representing perfect strong scaling.\label{fig:strong_scaling}}
\end{figure}

The strong scaling test we perform uses a grid setup similar to what we 
use for well-resolved binary simulations. With a single refined level, 
we have approximately $5 \times 10^7$ zones spread over $\sim 1000$ grids.
On a second refined level, there are $2 \times 10^9$ zones spread over 
$\sim 16000$ grids. We run a scaling test for both cases, with the 
highest processor count in each case chosen so that the number of 
MPI tasks is similar to the number of grids. There are no relative gains 
to be achieved from further parallelism. The results are found in 
\autoref{fig:strong_scaling}. We find excellent scaling for low numbers 
of processors, though the long wall time per timestep required for the 
smallest processor counts makes such a simulation infeasible in practice.
The best tradeoff between reasonable wall time and parallel efficiency 
seems to be when there are approximately 2--4 grids per processor. The 
scaling behavior worsens at the highest processor counts, but this is 
an expected consequence of processors becoming work-starved. In general 
we find very good strong scaling behavior in the regime we are presently 
interested in, simulating the early phases of a simulation at moderate 
resolution. The strong scaling behavior is acceptable, though not perfect, 
at very large processor counts when self-gravity is considered.

>>>>

To examine the parallel performance of \castro, we performed both
strong scaling and weak scaling tests on the Blue Waters machine at
the National Center for Supercomputing Applications. For the weak
scaling test, whose results are shown in \autoref{fig:weak_scaling},
we ran a uniform grid binary white dwarf simulation for resolutions of
$128^3$ zones through $2048^3$ zones. The number of processors was
scaled with the number of zones so that each processor had the same
amount of work; the smallest test used 8 processors and the largest
used 32,768 (note that the number of processor cores on a Blue Waters
node is twice the number of floating point units on that node). The
test was run for 10 timesteps, with each timestep including two
Poisson solves and a hydrodynamics update (though for a uniform grid
calculation we generally do not need to perform any multigrid
iterations for the first Poisson solve in a timestep, since the
density distribution has not changed since the end of the last
timestep). We disabled plotfile and checkpoint writing, as well as
calculation of diagnostic information (the latter can contribute to a
significant fraction of the run time at large processor counts if
computed every timestep). We computed the median wall time required
per time step for each simulation, and then normalized this to the
median time per timestep for the smallest simulation. We find
excellent weak scaling through 4,096 processors. At the largest run,
the simulation time required is slightly than 1.5 times the amount
required for the smallest simulation.  This is due entirely to the
increased cost of the multigrid Poisson solve in each timestep and
this cannot be mitigated except by improving communication or
computation efficiency in the multigrid solver. We observe that this
weak scaling behavior with Poisson gravity is a significant
improvement over the results presented in the first \castro\ paper.

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{plots/strong_scaling}
  \caption{\castro\ strong scaling test performed on the Blue Waters machine at
    NCSA. The vertical axis measures the median time per timestep, and the 
    horizontal axis measures the number of processors in the simulation. Data 
    points are normalized to the time per timestep for the smallest number 
    of processors. The green circles show the data for a simulation with one 
    AMR level (a single uniform grid), the blue diamonds show the data for a simulation with two AMR
    levels (one coarse and one fine), while the red circles show the data 
    for a test with three AMR levels (one coarse and two fine). The fine levels 
    increase the resolution only in the regions around the stars. For each case 
    we draw a solid curve representing perfect strong scaling.\label{fig:strong_scaling}}
\end{figure}

The strong scaling test we perform uses a grid setup similar to what we 
use for well-resolved binary simulations. With only a uniform coarse grid, 
there are approximately $2 \times 10^7$ zones. With a single refined level, 
we have approximately $2 \times 10^8$ zones, typically spread over $\sim 2000$ grids.
On a second refined level, there are a similar number of zones and grids 
(the volume covered by this level is smaller, which offsets the greater resolution).
We run a scaling test for all three cases, with the 
highest processor count in each case chosen so that the number of 
MPI tasks is similar to the number of grids. There are no gains 
to be achieved from further parallelism. The results are found in 
\autoref{fig:strong_scaling}. We find excellent scaling for low to moderate numbers 
of processors. Parallel efficiency is well maintained when there are 
at least 2 grids per processor. The 
scaling behavior worsens at the highest processor counts, but this is 
an expected consequence of processors becoming work-starved. At the highest 
processor count in this test, there is approximately only one grid per processor. In general 
we find very good strong scaling behavior in the regime we are presently 
interested in, simulating the early phases of a simulation at moderate 
resolution. The strong scaling behavior is acceptable, though not perfect, 
at very large processor counts when self-gravity is considered.
